Enhanced Material Point Method to face dynamic problems: Local-maximum entropy approximation and explicit predictor-corrector scheme\\ \PNA{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}

### 2020-04-28 LME-MPM.tex
Finished PNA

\newglossaryentry{neumann-boundary}
{
  name=$ \Gamma_n$,
  description={Natural or Nemann boundary conditions over $\partial\varOmega$}
}

@@

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers has proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing\PNA{somehow} the accuracy.\MMP{One of the first attempts of it was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} to downsize this spurious noises add a nonvoiscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. Within the explicit
time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort.} In this paper a less time consuming and high efficient explicit predictor-corrector integration method has been
proposed. It consists in an accommodation of the traditional \acrsull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and its good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

@@

\section{Conclusions}
\label{sec:conclusions}
We have developed a novel time integration scheme for
\acrshort{mpm}. The \acrshort{npc} arise as a highly efficient
alternative for demanding dynamic problems like coupled $u-p_w$
without appeal to expensive implicit time integration algorithms. Also
the procedure employed to design the \acrshort{npc} algorithm open the
door to revisit a huge variety of time integration schemes developed
originally for \acrshort{fem}, which can be rearranged to
\acrshort{mpm} framework with some modifications. Anyway, further
research should be done to elucidate the formal comprehension of the
algorithm good performance. This paper also enhances the skills of
local \textit{max-ent} approximation scheme as a robust and versatile
tool in the \acrshort{mpm} framework. It also come up as a promising
alternative to other approximation techniques developed for the
\acrshort{mpm} to overcome grid crossing limitations and to avoid the
constriction of the \acrshort{ugimp} of a regular mesh or a high
density of particles per cell. Future research of the group will be on
the employ of this scheme to strengthen the localization capabilities of
\acrshort{mpm} for viscoplastic materials. Finally we remark on the
possibility of adapting the function $\beta(\vect{x})$ through the finite strain
tensor in order to guide the shape function in the principal strain
direction to way select the most suitable set of nodes. Other
research line is the possibility to adapt the value of $\beta$ to solve the equations
\acrshort{fem}-like of meshfree-like depending of how behaves the
region, this could be extremely useful to reproduce such challenging
scenarios as simulating simultaneously both initialization and
propagation stages of fast landslides.
% 
\section*{Conflict of interest}
%
The authors declare that they have no conflict of interest.

>>>>

\newglossaryentry{neumann-boundary}
{
  name=$ \Gamma_n$,
  description={Natural or Neumann boundary conditions over $\partial\varOmega$}
}

@@

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers have proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing\PNA{somehow} the accuracy.\MMP{One of the first attempts was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} downsized these spurious noises by adding a non-viscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. Within the explicit
time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort.} In this paper a less time consuming and high efficient explicit predictor-corrector integration method has been
proposed. It consists of an accommodation of the traditional \acrsull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and its good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

@@

\section{Conclusions}
\label{sec:conclusions}
\PNA{The proposed research has dived into the performance of the \acrfull{mpm} within dynamic problems of very fast loading rate. The main goal, which was the enhancement of such a methodology with different time and spatial discretization tools in order to stabilize the solution, has been satisfactory fulfilled.}

\PNA{On the one hand,} a novel time integration scheme for
\acrshort{mpm} has been carried out. The \acrshort{npc} arises as a highly efficient
alternative for demanding dynamic problems \PNA{with high frequency of the applied loads. In the literature, several approaches of the \acrshort{npc} are proposed in order to compete on accuracy with expensive implicit time integration algorithms.} \PNA{The main achievement of the proposed time discretization has been observed within the Dyka's bar \ref{sec:dyka-bar}, obtaining lower values of the error than the ones obtained with the \acrfull{fe} scheme.}
Also,
the procedure employed to design the \acrshort{npc} algorithm open the
door to revisit a huge variety of time integration schemes developed
originally for \acrshort{fem}, which can be rearranged to the
\acrshort{mpm} framework with the required modifications. \PNA{As well as the implementation of enhanced time integration schemes, further
research may be done in the comprehension of the
good performance of the algorithm within non-linear solutions, such as the coupling between different phases or the non-linear constitutive behaviour of the materials: the employment of this scheme to strengthen the localization capabilities of \acrshort{mpm} for viscoplastic materials is on the scope of this research line. The comparison with traditional time schemes in the non-linear field will improve the formal understanding of the proposed methodology.}

\PNA{On the other hand, the spatial discretization of the \acrshort{mpm} is enhanced thanks to the powerful skills of the \acrfull{lme} approximation scheme, which has been validated as a robust and versatile
tool in the \acrshort{mpm} framework.} It also comes up as a promising
alternative to other approximation techniques\PNA{, developed within the
\acrshort{mpm} field, in order }to overcome grid crossing limitations and to avoid the
constriction of the \acrshort{ugimp} of a regular mesh or a high
density of particles per cell. \PNA{Several research lines of the \acrshort{lme} scheme dive into the improvement of the methodology, focusing in the optimization of the calculation of the shape function and the possibility of adapting the parameter $\beta(\vect{x})$ through the finite strain
tensor in order to guide the shape function in the principal strain
direction to select the most suitable set of nodes adapted to the strain, as well as the possibility of adapting} the value of $\beta$ to solve the equations
\acrshort{fem}-like of meshfree-like depending of how behaves the
region. \PNA{Together with the implementation of non-linear behaviours, previously mentioned, challenging
scenarios could be reproduced, being the main purpose of the research line the simultaneous simulation of both initialization and
propagation stages of fast landslides.}

% 
\section*{Conflict of interest}
%
The authors declare that they have no conflict of interest.
